---
title: "Members & Workstreams"
---

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(cowplot)
library(magick)
library(gridExtra)
library(readxl)
library(dplyr)
```

```{r read-team-info, echo=FALSE}
team_info_df <- read_xlsx("~/pharmaR.github.io/content/meet-the-team/teaminfo.xlsx")

largest_ws <- names(which.max(table(team_info_df$workstream)))

largest_ws_size <- length(which(team_info_df$workstream == largest_ws))
```

```{r photo-row-func, echo=FALSE}
photo_row_function <- function(filtered_df) {
  list <- list()
  for (person in 1:nrow(filtered_df)) {
  img_path <- paste0("~/pharmaR.github.io/content/meet-the-team/", filtered_df$photo_file_name[person])  
  p <- 
    ggdraw() + 
    draw_image(img_path, scale = .9)
  
  p <- ggdraw(add_sub(p, filtered_df$name[person], fontfamily = "helvetica", fontface = "bold", size = 10))
  
  list[[person]] <- ggdraw(add_sub(p, filtered_df$role[person], fontfamily = "helvetica", size = 10))
  }
  
  if (length(list) < largest_ws_size){
    img_path <- paste0("~/pharmaR.github.io/content/meet-the-team/blank.jpg")
    p <- 
      ggdraw() + 
      draw_image(img_path, scale = .9)
    
    p <- ggdraw(add_sub(p, " ", fontfamily = "helvetica", fontface = "bold", size = 10))
    
    list[[largest_ws_size]] <- ggdraw(add_sub(p, " ", fontfamily = "helvetica", size = 10))
    
  }

grid.arrange(grobs=c(list), nrow = 1, padding = unit(1, "cm"))
}
```

## Team Structure

The R Validation Hub is an R Consortium Working Group. The R Consortium supports several related working groups, and these working groups have a similar overall objective: to support the use of R within the biopharmaceutical industry. Furthermore, the working groups are complementary; each has its own distinct deliverables and many of the R Validation Hub's members are also members of the other working groups.

Within the R Validation Hub, we have many workstreams that focus on their own respective activities. Find each workstream's duties below as well as the members that support them.

### Strategic Workstream

The **Strategic Workstream** has the responsibility of aligning on longer-term strategic goals. The intention is that this workstream is to become a floating set of representatives from each of our workstreams.

```{r chair-rconsortium, echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 3}
chair_consortium_df <-
  team_info_df %>%
  filter(workstream %in% c("chair", "r consortium")) %>%
  arrange(priority, name)

photo_row_function(chair_consortium_df)
```

### Communications Workstream

The **Communications Workstream** focuses on how we build connections across the R world, specifically with our neighboring initiatives: the R Consortium, PhUSE, PSI AIMS, the R Submissions Working Group, and ROpenSci. This working group also supports the effort to make the R Validation Hub more intentional with how we organize ourselves and look into best means to disseminate information.

```{r comm-ws, echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 3}
comm_df <-
  team_info_df %>%
  filter(workstream == "communications") %>%
  arrange(priority, name)

photo_row_function(comm_df)
```

### Repositories Workstream

The **Repositories Workstream** supports a transparent, cross-industry approach of establishing and maintaining a repository of validated R packages.

```{r repo-ws, echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 3}
repo_df <-
  team_info_df %>%
  filter(workstream == "repositories") %>%
  arrange(priority, name)

photo_row_function(repo_df)
```

### {riskassessment} Workstream

[Text about workstream]

```{r riskassessment-ws, echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 3}
riskassessment_df <-
  team_info_df %>%
  filter(workstream == "riskassessment") %>%
  arrange(priority, name)

photo_row_function(riskassessment_df)
```

### {riskmetric} Workstream

[Text about workstream]

```{r riskmetric-ws, echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 3}
riskmetric_df <-
  team_info_df %>%
  filter(workstream == "riskmetric") %>%
  arrange(priority, name)

photo_row_function(riskmetric_df)
```

### Previous Members
