---
title: "Members & Workstreams"
---

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(cowplot)
library(magick)
library(gridExtra)
library(readxl)
library(dplyr)
```

```{r read-team-info, echo=FALSE}
team_info_df <- read_xlsx("content/meet-the-team/teaminfo.xlsx")
largest_ws <- names(which.max(table(team_info_df$workstream)))
largest_ws_size <- length(which(team_info_df$workstream == largest_ws))
```

```{r photo-row-func, echo=FALSE}
photo_row_function <- function(filtered_df) {
  grobs <- list()

  for (person in seq_len(nrow(filtered_df))) {
    img_path <- paste0(
      "content/meet-the-team/",
      filtered_df$photo_file_name[person]
    )

    p <-  ggdraw() + draw_image(img_path, scale = .9)
    p <- ggdraw(add_sub(
      p,
      filtered_df$name[person],
      fontface = "bold",
      size = 10
    ))

    grobs[[person]] <- ggdraw(add_sub(
      p,
      filtered_df$role[person],
      size = 10
    ))
  }

  if (length(grobs) < largest_ws_size) {
    img_path <- paste0("content/meet-the-team/blank.jpg")
    p <- ggdraw() + draw_image(img_path, scale = 0.9)
    p <- ggdraw(add_sub(
      p,
      " ",
      fontface = "bold",
      size = 10
    ))

    grobs[[largest_ws_size]] <- ggdraw(add_sub(
      p,
      " ",
      size = 10
    ))
  }

  grobs <- Filter(Negate(is.null), grobs)
  grid.arrange(grobs = grobs, nrow = 1, padding = unit(1, "cm"))
}
```

### Our Structure

The R Validation Hub champions a level structure, comprised of many workstreams that contribute toward their focus areas and activities. Find each workstream's duties below as well as the members that support them.

*If you wish to contribute to any of our efforts, please visit our [Contribute](https://www.pharmar.org/contribute/) page.*

### Strategic Workstream

The **Strategic Workstream** has the responsibility of aligning on longer-term strategic goals. The intention is that this workstream is to become a floating set of representatives from each of our workstreams.

```{r chair-rconsortium, echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 3}
chair_consortium_df <-
  team_info_df %>%
  filter(workstream %in% c("chair", "r consortium")) %>%
  arrange(priority, name)

photo_row_function(chair_consortium_df)
```

### Communications Workstream

The **Communications Workstream** focuses on how we build connections across the R world, specifically with our neighboring initiatives: the R Consortium, PhUSE, PSI AIMS, the R Submissions Working Group, and ROpenSci. This working group also supports the effort to make the R Validation Hub more intentional with how we organize ourselves and look into best means to disseminate information.

```{r comm-ws, echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 3}
comm_df <-
  team_info_df %>%
  filter(workstream == "communications") %>%
  arrange(priority, name)

photo_row_function(comm_df)
```

### Repositories Workstream

The **Repositories Workstream** supports a transparent, cross-industry approach of establishing and maintaining a repository of validated R packages.

```{r repo-ws, echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 3}
repo_df <-
  team_info_df %>%
  filter(workstream == "repositories") %>%
  arrange(priority, name)

photo_row_function(repo_df)
```

### {riskassessment} Application Workstream

The **{riskassessment} Workstream** leads the development of the risk assessment Shiny app (an extension of the {riskmetric} package) that provides a graphical interface to the package's functionality.

```{r riskassessment-ws, echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 3}
riskassessment_df <-
  team_info_df %>%
  filter(workstream == "riskassessment") %>%
  arrange(priority, name)

photo_row_function(riskassessment_df)
```

### {riskmetric} Workstream

The **{riskmetric} Workstream** leads the development of the {riskmetric} package that supports the idea of risk-based evaluation of R packages.

```{r riskmetric-ws, echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 3}
riskmetric_df <-
  team_info_df %>%
  filter(workstream == "riskmetric") %>%
  arrange(priority, name)

photo_row_function(riskmetric_df)
```

### Previous Members

*Andy Nicholls* (Chair 2018-2022)

*Lyn Taylor*, *Paulo Bargo*, *Marly Gotti*, *Yilong Zhang*, *Keaven Anderson*, *Min Lee*, *Parker Sims*
