---
title: "Participating Organizations"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(cowplot)
library(magick)
library(gridExtra)
library(readxl)
library(dplyr)
```

```{r get_photos, echo=FALSE}
list_of_logos <- sort(list.files("~/pharmaR.github.io/content/participating_orgs_photos/"))
logo_path_list <- c()
for (logo in 1:length(list_of_logos)) {
  logo_path <- paste0("~/pharmaR.github.io/content/participating_orgs_photos/", list_of_logos[logo])
  logo_path_list <- c(logo_path_list, logo_path)
}
```

The R Validation Hub is comprised of participants from across the pharmaceutical industry. Participants contribute to the effort through our regular group meetings, as well as the various workstreams that make up the project. Feel free to explore [contribution opportunities](https://www.pharmar.org/contribute) and subscribe to [our mailing list](https://lists.r-consortium.org/g/RConsortium-Validation-Hub/) to stay up-to-date on our progress.

*If you are a member and your organization does not appear on this list, please let us know!*

```{r photo-row-func, echo=FALSE, fig.width = 11}
photo_row_function_logos <- function(path_list) {
  list <- list()
  for (org in 1:length(logo_path_list)) {
    logo_path <- logo_path_list[org]  
    p <- 
      ggdraw() + 
      draw_image(logo_path, scale = .95)
    
    list[[org]] <- p
  }

grid.arrange(grobs=c(list), nrow = 6, ncol = 9)
}

photo_row_function_logos(logo_path_list)
```
